Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-10-01 writeup/hat-vldb.tex
Intro tweaks for CR

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
%\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}
\usepackage[hyphens]{url}

@@

@@

\title{Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi}
{\author{Peter Bailis{\fontsize{12}{14}$^\dagger$}, Aaron Davidson{\fontsize{12}{14}$^\dagger$}, Alan Fekete{\fontsize{12}{14}$^\diamond$}, Ali Ghodsi{\fontsize{12}{14}$^{\dagger,\ddagger}$}, Joseph M. Hellerstein{\fontsize{12}{14}$^\dagger$}, Ion Stoica{\fontsize{12}{14}$^\dagger$}\\{\affaddr{\fontsize{12}{14}$^\dagger$}\hspace{.5mm}UC Berkeley\hspace{4mm}{\fontsize{12}{14}$^\diamond$}\hspace{.5mm}University of Sydney\hspace{4mm}{\fontsize{12}{14}$^\ddagger$}\hspace{.5mm}KTH/Royal Institute of Technology}}}
\maketitle

>>>>

\documentclass{vldb}
\let\proof\relax
\let\endproof\relax
\usepackage{draftwatermarkbottom, draftwatermarktop}
\usepackage{scrextend, MnSymbol}
\usepackage[hyphens]{url}

@@

\newenvironment{introenumerate}
{

   \vspace{-.5em}
   \newcounter{qdcounter}
    \begin{list}{\arabic{qdcounter}.~}{\usecounter{qdcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

@@

\title{Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi}
{\author{Peter Bailis, Aaron Davidson, Alan Fekete{\fontsize{12}{14}$^\dagger$}, Ali Ghodsi, Joseph M. Hellerstein, Ion Stoica \\{\affaddr{UC Berkeley and {\fontsize{12}{14}$^\dagger$}University of Sydney}}}}
\maketitle

### 2013-10-01 writeup/intro.tex
Intro tweaks for CR

At a high level, the virtues of HATs are guaranteed responses from any
replica, low latency, and a range of semantic guarantees including
several whose usefulness is widely acepted such as Read
Committed. However, highly available systems are fundamentally unable
to prevent concurrent updates to shared data items and cannot provide
recency guarantees for reads. To understand when these virtues and
limitations are relevant in practice, we survey both practitioner
accounts and academic literature, perform experimental analysis on
modern cloud infrastructure, and analyze representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, particularly for programs with
non-monotonic logic, HATs can fall short, requiring
sometimes-unavailable mechanisms that are more
coordination-intensive. Overall, this paper quantifies the benefits
that Highly Available Transactions offer as well as the necessary
restrictions that they place on applications.

>>>>

At a high level, the virtues of HATs are guaranteed responses from any
replica, low latency, and a range of semantic guarantees including
several whose usefulness is widely acepted such as Read
Committed. However, highly available systems are fundamentally unable
to prevent concurrent updates to shared data items and cannot provide
recency guarantees for reads. To understand when these virtues and
limitations are relevant in practice, we survey both practitioner
accounts and academic literature, perform experimental analysis on
modern cloud infrastructure, and analyze representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, particularly for programs with
non-monotonic logic or that are otherwise concurrency-sensitive, HATs
can fall short, requiring non-HA coordination.

We recognize that the large variety of ACID isolation levels and
distributed consistency models (and therefore in our taxonomization)
can be confusing and that the subtle distinctions between them may
appear to be primarily of academic concern. Accordingly, we offer the
following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

\item Most of these ``weak isolation'' models are achievable without
  sacrificing high availability if implemented correctly. However,
  none of these models prevents concurrent modifications.

\item In addition to providing a guaranteed response and horizontal
  scale-out, these HAT models allow for one to three order of
  magnitude lower latencies on current infrastructure.

\item For correct behavior, most applications will require a
  combination of HAT and (ideally sparing use of) non-HAT isolation levels;
  future database designers should act accordingly.
\end{introenumerate}
